---
title: "1. 2024 Rapid Assessment Plots"
author: "Christina Fossum"
date: "2025-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

February 11, 2025
Christina Fossum
Purpose: QAQC Functions for ROMO PIPO (FMH) Plot Data

# Step 1. Load packages and functions
```{r}
#Load packages
library(tidyverse)
library(dplyr)
library(writexl)
library(knitr)
library(EnvStats)

# Create Blank error log
errors_blank <- data.frame("SavedQuery" = "", "MacroPlot.Name" = "", "Sample.Event.Date" = "", "Error" = "No Error", "Comment" = "")

# QAQC Function
qaqc <- function(data, query, query_message, values_check, Comment) {
  errors <- data %>%   
    mutate("SavedQuery" = query, 
           "Error" = paste(query_message, "=", values_data)) %>%
    filter(!(values_check) %>% replace_na(TRUE)) %>%   
    select("SavedQuery", "MacroPlot.Name", "Sample.Event.Date", "Error", "Comment")
  if (nrow(errors) == 0) {   
    errors <- errors_blank %>%     
      mutate(SavedQuery = query) 
  } else {   
    errors <- errors
  }
}


# Header loading function
load_headers <- function(directory_path, pattern) {
  list_files <- list.files(path = directory_path, pattern = pattern, full.names = TRUE)
  data_list <- lapply(list_files, function(file) {
    file_data <- read.csv(file) %>% select(1:10) %>% filter(Visited == "True")
    file_data <- file_data %>% mutate(Comment = "")
  })
  return(do.call(rbind, data_list))
}



# Data Loading Function
load_data <- function(directory_path, pattern) {
  list_files <- list.files(path = directory_path, pattern = pattern, full.names = TRUE)
  data_list <- lapply(list_files, function(file) {
    file_data <- read.csv(file, skip = 3) 
    file_data <- file_data %>% mutate(Comment = "")
  })
  return(do.call(rbind, data_list))
}

```


# Step 2: Load Data
```{r}

#Set Working Directory to QAQC Folder: Session -> Set Working Directory -> Choose Directory

# Set "directory_path" to source file location
directory_path <- "o:/resmgmt/Fire Management/Fire Effects/_QAQC/FFI data - WORKING/FMH_PIPO_2024/"


# load data (Surface Fuels - Fine)
fuelsfine_df <- load_data(directory_path, "_Surface Fuels - Fine.csv$")
fuelsfine_headers <-load_headers(directory_path, "_Surface Fuels - Fine.csv$")

# load data (Surface Fuels - 1000Hr)
fuels1000_df <- load_data(directory_path, "_Surface Fuels - 1000Hr.csv$")
fuels1000_headers <-load_headers(directory_path, "_Surface Fuels - 1000Hr.csv$")

# load data (Surface Fuels - Duff/Litter)
fuelsDL_df <- load_data(directory_path, "_Surface Fuels - Duff_Litter.csv$")
fuelsDL_headers <-load_headers(directory_path, "_Surface Fuels - Duff_Litter.csv$")

# load data (Trees - Individuals)
trees_df <- load_data(directory_path, "_Trees - Individuals \\(metric\\)\\.csv$")

list <- list.files(path = directory_path, pattern = "_Trees - Individuals \\(metric\\)\\.csv$", full.names = TRUE)
trees_headers <- lapply(list, function(file) read.csv(file, nrows = 1))
trees_headers <- do.call(rbind, trees_headers)

# load data (Cover - Points)
cover_df <- load_data(directory_path, "_Cover - Points \\(metric\\)\\.csv$")
cover_headers <-load_headers(directory_path, "_Cover - Points \\(metric\\)\\.csv$")

# load data (Post Burn Severity)
BS_df <- load_data(directory_path, "_Post Burn Severity \\(metric\\)\\.csv$")
BS_headers <-load_headers(directory_path, "_Post Burn Severity \\(metric\\)\\.csv$")

# load data (shrub belt transects)
shrub_df <- load_data(directory_path, "_Density - Belts \\(metric\\)\\.csv$")
shrub_headers <-load_headers(directory_path, "_Density - Belts \\(metric\\)\\.csv$")

# load data (Seedlings)
seedling_df <- load_data(directory_path, "_Density - Quadrats \\(metric\\)\\.csv$")
seedling_headers <-load_headers(directory_path, "_Density - Quadrats \\(metric\\)\\.csv$")



```


##Step 3: Data Checks                                      


# Surface Fuels - Fine (metric)

Checks:
1. Number of Transects = 4
2. Transect Length = 6 for 1hr, 6 for 10hr, 12 for 100hr
3. Reasonable Azimuth
4. FWD fuel constant = Default
4. 4x transect data per plot
5. Outliers flagged for 1hr, 10hr, and 100hr tally


```{r}

# QAQC Checks (for Fine Fuels) 

# 1. Number of Transects = 4
values_data <- fuelsfine_headers$Num..Transects #this is the column we are checking
values_valid <- 4
values_check <- values_data == values_valid
errors_FuelsFine <- qaqc(fuelsfine_headers, "Fine Fuels NumTran", "Number of Transects", values_check)

# 2. Transect Length = 6ft for 1hr & 10hr, 12ft for 100hr
values_data1 <- fuelsfine_headers$X1.hr.Tran..Len. #1-hr transect length
values_data10 <- fuelsfine_headers$X10.hr.Tran..Len. #10-hr transect length
values_data100 <- fuelsfine_headers$X100.hr.Tran..Len. #100-hr transect length
values_valid <- 6
values_valid2 <- 12
values_check1 <- values_data1 == values_valid
values_check10 <- values_data10 == values_valid
values_check100 <- values_data100 == values_valid2
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_headers, "fine fuels 1hr TranLength", "1hr Transect Length", values_check1))
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_headers, "fine fuels 10hr TranLength", "10hr Transect Length", values_check10))
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_headers, "fine fuels 100hr TranLength", "100hr Transect Length", values_check100))

# 3. Azimuth Checks
values_data <- fuelsfine_df$Azimuth
values_valid <- seq(0, 359, by = 1)
values_check <- values_data %in% values_valid
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_df, "fine fuels azimuth", "Azimuth", values_check))

# 4. Check Fuel Constant
values_data <- fuelsfine_df$FWDFuConSt
values_valid <- "Default" 
values_check <- values_data == values_valid
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_df, "Fuel Fine FuelConst", "Fuel Constant", values_check))


# 5. Hits Check
fuelsfine_df <- fuelsfine_df %>% mutate(hits = 1) %>%
  group_by(MacroPlot.Name, Sample.Event.Date) %>%
  mutate(sumhits = sum(hits)) %>% ungroup()
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_df, "Fuel Fine Hits", "Number of Hits", fuelsfine_df$sumhits == 4))

# 6. Outlier Detection for 1hr, 10hr, 100hr
test1hr <- rosnerTest(fuelsfine_df$X1.Hr.Count)
test10hr <- rosnerTest(fuelsfine_df$X10.Hr.Count)
test100hr <- rosnerTest(fuelsfine_df$X100.Hr.Count)
maxoutlier1hr <- max(test1hr$all.stats$Value[test1hr$all.stats$Outlier])
maxoutlier10hr <-  max(test10hr$all.stats$Value[test10hr$all.stats$Outlier])
maxoutlier100hr <- max(test100hr$all.stats$Value[test100hr$all.stats$Outlier])
values_check_1hr <- fuelsfine_df$X1.Hr.Count < maxoutlier1hr
values_check_10hr <- fuelsfine_df$X10.Hr.Count < maxoutlier10hr
values_check_100hr <- fuelsfine_df$X100.Hr.Count < maxoutlier100hr

errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_df, "Fine Fuels x1hr value", "One Hour Count Outlier", values_check_1hr))
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_df, "Fine Fuels x10hr value", "10 Hour Count Outlier", values_check_10hr))
errors_FuelsFine <- rbind(errors_FuelsFine, qaqc(fuelsfine_df, "Fine Fuels x100hr value", "100 Hour Count Outlier", values_check_100hr))


```

# Surface Fuels - 1000Hr (metric)

Checks: 
1. Number of transects = 4
2. Transect length = 100ft
3. Correct CWD Fuel Constant
4. Decay Class is either 3 or 4
5. Outliers flagged for diameter

```{r}

# QAQC Checks 

# 1. Number of Transects = 4
values_data <- fuels1000_headers$Num..Transects
values_valid <- 4
values_check <-  values_data == values_valid
errors_Fuels1000 <- qaqc(fuels1000_headers, "1000hr Fuels NumTran", "Number of Transects", values_check)

# 2. Transect Length = 100
values_data <- fuels1000_headers$Tran..Length
values_valid <- 100
values_check <-  values_data == values_valid
errors_Fuels1000 <- rbind(errors_Fuels1000, qaqc(fuels1000_headers, "x1000 Fuels TranLen", "Transect Length", values_check))

# 3. Check Fuel Constant
values_data <- fuels1000_df$CWDFuConSt
values_valid <- "Default"
values_check <- values_data == values_valid
errors_Fuels1000 <- rbind(errors_Fuels1000, qaqc(fuels1000_df, "x1000 CWD FuelConst" , "Fuel Constant", values_check))

# 4. Decay class 3 or 4
valid_decay_classes <- c(3, 4)
values_check <- fuels1000_df$Decay.Class %in% valid_decay_classes
errors_Fuels1000 <- rbind(errors_Fuels1000, qaqc(fuels1000_df, "x1000 decay class", "decay class 3 or 4", values_check))

# 5. Outliers flagged for diameter
test1000hr <- rosnerTest(fuels1000_df$Diameter)
maxoutlier <- max(test1000hr$all.stats$Value[test1000hr$all.stats$Outlier])
values_check <- fuels1000_df$Diameter < maxoutlier
errors_Fuels1000 <- rbind(errors_Fuels1000, qaqc(fuels1000_df, "x1000 Dia value", "Diameter outlier", values_check))


```

# Surface Fuels - Duff/Litter (metric)

Checks:
1. # of transects = 4
2. Correct Sample Locations: 1, 5, 10, 15, 20, 25, 30, 35, 40, 45
3. # of samples per transect = 10
4. Correct DL Fuel Constant
5. Outlier values flagged for duff, & litter


```{r}


# QAQC Checks 

# 1. Number of Transects = 4
values_data <- fuelsDL_headers$Num..Transects
values_valid <- 4
values_check <- values_data == values_valid
errors_FuelsDL<- qaqc(fuelsDL_headers, "DL NumTran", fuelsDL_headers$Num..Transects, values_check)

# 2. Correct Sample Locations
values_data <- fuelsDL_df$Samp..Loc.
values_valid <- c(1, 5, 10, 15, 20, 25, 30, 35, 40, 45)
values_check <- values_data %in% values_valid
errors_FuelsDL <- rbind(errors_FuelsDL, qaqc(fuelsDL_df, "Sample Locations", "Tape", values_check))


# 3. # samples per transect = 10
fuelsDL_df <- fuelsDL_df %>% mutate(hits = 1) %>% group_by(MacroPlot.Name, Sample.Event.Date, Transect) %>%
  mutate(sumhits = sum(hits)) %>% ungroup()
values_data <- fuelsDL_df$sumhits
values_valid <- 10
values_check <- values_data == values_valid
errors_FuelsDL <- rbind(errors_FuelsDL, qaqc(fuelsDL_df, "LitterDuff Hits Per Transect", "Hits Per Transect", values_check))

# 4. Correct DL Fuel Constant
values_data <- fuelsDL_df$DLFuConSt
values_valid <- "Default"
values_check <- values_data == values_valid
errors_FuelsDL <- rbind(errors_FuelsDL, qaqc(fuelsDL_df, "DL FuelConst" , fuelsDL_df$DLFuConSt, values_check))

# 5. OUtliers flagged

#Litter
fuelsDL_df$Litter.Depth[is.na(fuelsDL_df$Litter.Depth)] <- 0
temp_data <- na.omit(fuelsDL_df$Litter.Depth)
test <- rosnerTest(temp_data)
outliers <- test$all.stats$Value[test$all.stats$Outlier]
max_val <- min(outliers) - 0.1
values_data <- fuelsDL_df$Litter.Depth
values_check <- values_data < max_val | is.na(values_data)
errors_FuelsDL <- rbind(errors_FuelsDL, qaqc(fuelsDL_df, "Litter Outliers", "Litter Depth", values_check))

#Duff
fuelsDL_df$Duff.Depth[is.na(fuelsDL_df$Duff.Depth)] <- 0
temp_data <- na.omit(fuelsDL_df$Duff.Depth)
test <- rosnerTest(temp_data)
outliers <- test$all.stats$Value[test$all.stats$Outlier]
max_val <- min(outliers) - 0.1
values_data <- fuelsDL_df$Duff.Depth
values_check <- values_data < max_val | is.na(values_data)
errors_FuelsDL <- rbind(errors_FuelsDL, qaqc(fuelsDL_df, "Duff Outliers", "Duff Depth", values_check))



```

# Trees - Individuals (metric)

Checks: 
1. Subplot = 1 or 2
2. Quarter = 1, 2, 3, or 4
3. Subplot 1 if DBH >= 15; subplot 2 if DBH < 15
4. Species = PIPO
5. Outlier values flagged for DBH, Char Height, Scorch Height, Scorch %
6. Correct Crown Class for snags
7. Correct Plot Area
8. Correct Snag Plot Area
9. Correct Break Pt. Diameter

```{r}

# QAQC Checks

# 1. Subplot = 1 or 2
values_data <- trees_df$Subplot.Frac.
valid_values <- c(1, 2)
values_check <- values_data %in% valid_values
errors_Trees<- qaqc(trees_df, "Subplot Fraction", trees_df$Subplot.Frac., values_check)

# 2. Quarter 1, 2, 3, 4
values_data <- trees_df$Quarter
valid_values <- c(1, 2, 3, 4)
values_check <- values_data %in% valid_values
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "Tree Quarter", trees_df$Quarter, values_check))

# 3. Subplot 1 if DBH >= 15; subplot 2 if DBH < 15
values_data <- trees_df$Subplot.Frac.
values_valid <- case_when(
  is.na(trees_df$DBH) | is.na(trees_df$Subplot.Frac.) ~ FALSE,  # Flag NA values as invalid
  trees_df$DBH >= 15 & trees_df$Subplot.Frac. == 1 ~ TRUE,
  trees_df$DBH < 15 & trees_df$Subplot.Frac. == 2 ~ TRUE,
  TRUE ~ FALSE
)
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "Tree subplot correct", trees_df$DBH, values_valid))

# 4. Species = PIPO
values_check <- trees_df$Species == "PIPO"
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "Tree species", trees_df$Species, values_check))

# 5. Outlier values flagged for DBH, Char Height, Scorch Height, Scorch %

###DBH
outliers <- tryCatch({
  rosnerTest(na.omit(trees_df$DBH))$all.stats$Value
}, error = function(e) {
  return(NULL)
})
if (!is.null(outliers) && length(outliers) > 0) {
  min_outlier <- min(outliers) + 1
  max_outlier <- max(outliers) - 1
} else {
  min_outlier <- min(trees_df$DBH, na.rm = TRUE)
  max_outlier <- max(trees_df$DBH, na.rm = TRUE)
}
values_data <- trees_df$DBH
values_check <- trees_df$DBH > min_outlier & trees_df$DBH < max_outlier
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "DBH outliers", "DBH", values_check))
###Char Height
outliers <- tryCatch({
  rosnerTest(na.omit(trees_df$Char.Ht.))$all.stats$Value
}, error = function(e) {
  return(NULL)
})
if (!is.null(outliers) && length(outliers) > 0) {
  min_outlier <- min(outliers) + 1
  max_outlier <- max(outliers) - 1
} else {
  min_outlier <- min(trees_df$Char.Ht., na.rm = TRUE)
  max_outlier <- max(trees_df$Char.Ht., na.rm = TRUE)
}
values_data <- trees_df$Char.Ht.
values_check <- trees_df$Char.Ht. > min_outlier & trees_df$Char.Ht. < max_outlier
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "Char Ht outliers", "Char Ht", values_check))
###Scorch Height
outliers <- tryCatch({
  rosnerTest(na.omit(trees_df$Scorch.Ht.))$all.stats$Value
}, error = function(e) {
  return(NULL)
})
if (!is.null(outliers) && length(outliers) > 0) {
  min_outlier <- min(outliers) + 1
  max_outlier <- max(outliers) - 1
} else {
  min_outlier <- min(trees_df$Scorch.Ht., na.rm = TRUE)
  max_outlier <- max(trees_df$Scorch.Ht., na.rm = TRUE)
}
values_data <- trees_df$Scorch.Ht.
values_check <- trees_df$Scorch.Ht. > min_outlier & trees_df$Scorch.Ht. < max_outlier
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "Scorch Ht outliers", "Scorch Ht", values_check))
###Scorch Percent
outliers <- tryCatch({
  rosnerTest(na.omit(trees_df$Crown.Scorch..))$all.stats$Value
}, error = function(e) {
  return(NULL)
})
if (!is.null(outliers) && length(outliers) > 0) {
  min_outlier <- min(outliers) + 1
  max_outlier <- max(outliers) - 1
} else {
  min_outlier <- min(trees_df$Crown.Scorch.., na.rm = TRUE)
  max_outlier <- max(trees_df$Crown.Scorch.., na.rm = TRUE)
}
values_data <- trees_df$Crown.Scorch..
values_check <- trees_df$Crown.Scorch.. > min_outlier & trees_df$Crown.Scorch.. < max_outlier
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "Scorch % outliers", "Scorch %", values_check))

# 6. Correct Crown Class for Snags
values_valid <- case_when(trees_df$Status == "D" & trees_df$Crown.Class %in% c("BAD", "LBS", "RS") ~ TRUE,
                          trees_df$Status == "L" & trees_df$Crown.Class %in% c("I", "O", "SC", "D", "C") ~ TRUE,TRUE ~ FALSE)
errors_Trees <- rbind(errors_Trees, qaqc(trees_df, "Tree CPC", trees_df$Crown.Class, values_valid))

# 7. Correct Plot Area
values_data <- trees_headers$Plot.Area
values_valid <- 0.1
values_check <- values_data == values_valid
errors_Trees <- rbind(errors_Trees, qaqc(trees_headers, "Tree Plot Area", trees_headers$Plot.Area, values_check))

# 8. Correct Snag Plot Area
values_data <- trees_headers$Snag.Plot.Area
values_valid <- 0.1
values_check <- values_data == values_valid
errors_Trees <- rbind(errors_Trees, qaqc(trees_headers, "Snag Plot Area", trees_headers$Snag.Plot.Area, values_check))

# 9. Correct Break Pt. Diameter
values_data <- trees_headers$Break.Pnt..Dia.
values_valid <- 15
values_check <- values_data == values_valid
errors_Trees <- rbind(errors_Trees, qaqc(trees_headers, "Tree Break Pt", trees_headers$Break.Pnt..Dia., values_check))


```

# Cover - Points (metric)

Checks: 
1. # of transects = 2
2. Transect Length = 50m
3. Points per transect = 166
4. NumPts = 166
5. 1st hit has height value only if a plant (no other height values)
6. Outlier values flagged for plant height
7. Outlier values flagged for cover species

```{r}

# QAQC Checks 

# 1. # Transects = 2
values_data <- cover_headers$Num..Transects 
values_valid <- 2 
values_check <- values_data == values_valid
errors_cover <- qaqc(cover_headers, "Cover NumTran", cover_headers$Num..Transects, values_check)

# 2. Transect Length = 50m
values_data <- cover_headers$Tran..Length
values_valid <- 50
values_check <- values_data == values_valid
errors_cover <- rbind(errors_cover, qaqc(cover_headers, "Cover Tran Length", cover_headers$Tran..Length, values_check))
 
# 3. Points per transect = 166
values_data <- cover_headers$Num..Pts..Tran.
values_valid <- 166
values_check <- values_data == values_valid
errors_cover <- rbind(errors_cover, qaqc(cover_headers, "Cover pts per transect", cover_headers$Num..Pts..Tran., values_check))

# 4. NumPts = 166
cover_df <- cover_df %>% group_by(MacroPlot.Name, Sample.Event.Date, Transect) %>% mutate(NumPts = n_distinct(Point)) %>% ungroup()
values_data <- cover_df$NumPts
values_valid <- 166
values_check <- values_data == values_valid
errors_cover <- rbind(errors_cover, qaqc(cover_df, "cover obvs #", cover_df$NumPts, values_check))

# 5. 1st hit has height value only if a plant (no other height values)
values_valid <- case_when(
  cover_df$Order == 1 & !(cover_df$Species %in% c("LITT1", "WOOD1", "BARE1", "ROCK1", "MOSS1")) & cover_df$Height >= 0 ~ TRUE,
  cover_df$Order != 1 & is.na(cover_df$Height) ~ TRUE,
  cover_df$Order == 1 & cover_df$Species %in% c("LITT1", "WOOD1", "BARE1", "ROCK1", "MOSS1") & is.na(cover_df$Height) ~ TRUE,
  TRUE ~ FALSE)
errors_cover <- rbind(errors_cover, qaqc(cover_df, "cover height", cover_df$Point, values_valid))

# 6. Outlier values flagged for plant height
cover_df[is.na(cover_df)] <- 0
outliers <- rosnerTest(na.omit(cover_df$Height))$all.stats$Value # 1.9, 1.8 are outliers
max_val <- min(outliers) - 0.1
values_data <- cover_df$Height
valid_values <- c(seq(0, max_val, by = 0.01), NA)  
values_check <- cover_df$Height %in% valid_values
errors_cover <- rbind(errors_cover, qaqc(cover_df, "cover height outliers", cover_df$Height, values_check))

# 7. Outlier values flagged for cover species
cover_df <- cover_df %>% group_by(Species) %>% mutate(SpeciesFreq = n()) %>% ungroup()
valid_values <- case_when(cover_df$SpeciesFreq >= 2 ~ TRUE, TRUE ~ FALSE)
errors_cover <- rbind(errors_cover, qaqc(cover_df, "species outlier" , cover_df$Species, valid_values))

```

# Post Burn Severity

Checks: 
1. # of transects = 4
2. Transect length = 15.2m
3. Points per transect = 10
4. Each transect has 10 rows of data
5. Correct sample locations (1, 5, 10, 15, 20, 25, 30, 35, 40, 45)

```{r}

# QAQC Checks 

# 1. # Transects = 4
values_data <- BS_headers$Num..Transects
values_valid <- 4 
values_check <- values_data == values_valid
errors_BS <- qaqc(BS_headers, "BS NumTran", BS_headers$Num..Transects, values_check)

# 2. Transect Length = 15.2m
values_data <- BS_headers$Tran..Length
values_valid <- 15.2
values_check <- values_data == values_valid
errors_BS <- rbind(errors_BS, qaqc(BS_headers, "BS Tran Length", BS_headers$Tran..Length, values_check))
 
# 3. Points per transect = 10
values_data <- BS_headers$Num..Pts..Tran.
values_valid <- 10
values_check <- values_data == values_valid
errors_BS <- rbind(errors_BS, qaqc(BS_headers, "BS pts per transect", BS_headers$Num..Pts..Tran., values_check))

# 4. Each transect has 10 rows of data
BS_df <- BS_df %>% group_by(MacroPlot.Name, Sample.Event.Date, Transect) %>% mutate(Nrows = n()) %>% ungroup()
values_data <- BS_df$Nrows
values_valid <- 10
values_check <- values_valid == values_data
errors_BS <- rbind(errors_BS, qaqc(BS_df, "BS Nrows", "number of rows", values_check))

# 5. Correct sample locations (1, 5, 10, 15, 20, 25, 30, 35, 40, 45)
values_data <- BS_df$Tape
values_valid <- c(1, 5, 10, 15, 20, 25, 30, 35, 40, 45)
values_check <- values_valid == values_data
errors_BS <- rbind(errors_BS, qaqc(BS_df, "BS Sample Locations", "Sample Locations", values_check))


```

# Density - Shrub belts

Checks:
1. # of transects = 2
2. # of sub-belts = 10
3. Transect length = 50m
4. Transect Width = 2m
5. Transect Area = 100m
6. Sub-belt = 1, 2, 3, 4, 5, 6, 7, 8, 9, or 10
7. Transect = 1 or 2
8. Age Class = I or M
9. Status = L or D
10. Count = 1-50
11. Species = CEFE, PUTR2, CEVE, RICE, RUID, BERBE, CHVIL4

```{r}

# QAQC Checks

# 1. # Transects = 2
values_data <- shrub_headers$Num..Transects
values_valid <- 2
values_check <- values_data == values_valid
errors_shrubs <- qaqc(shrub_headers, "shrub NumTran", "Number of Transects", values_check)

# 2. # of sub-belts = 10
values_data <- shrub_headers$Num..Subbelts
values_valid <- 10
values_check <- values_data == values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_headers, "shrub NumSubbelts", "Subbelts per transect", values_check))

# 3. Transect length = 50m
values_data <- shrub_headers$Tran..Length
values_valid <- 50
values_check <- values_data == values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_headers, "shrub TranLength", "Transect Length", values_check))

# 4. Transect width = 2m
values_data <- shrub_headers$Tran..Width
values_valid <- 2
values_check <- values_data == values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_headers, "shrub TranWidth", "Transect Width", values_check))

# 5. Transect Area = 100m
values_data <- shrub_headers$Tran..Area
values_valid <- 100
values_check <- values_data == values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_headers, "shrub TranArea", "Transect area", values_check))

# 6.Sub-belt = 1, 2, 3, 4, 5, 6, 7, 8, 9, or 10
values_data <- shrub_df$Subbelt
values_valid <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
values_check <- values_data %in% values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_df, "shrub subbelt #", "subbelt", values_check))

# 7. Transect = 1 or 2
values_data <- shrub_df$Transect
values_valid <- c(1, 2)
values_check <- values_data %in% values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_df, "shrub transect #", "transect", values_check))

# 8. Age Class = I or M
values_data <- shrub_df$Age.Class
values_valid <- c("I", "M")
values_check <- values_data %in% values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_df, "shrub age class", "age class", values_check))

# 9. Status = L or D
values_data <- shrub_df$Status
values_valid <- c("L", "D")
values_check <- values_data %in% values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_df, "shrub status", "status", values_check))

# 10. Count = 1-50
values_data <- shrub_df$Count
values_valid <- seq(0, 50, by = 1)
values_check <- values_data %in% values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_df, "shrub count", "count", values_check))

# 11. Species = CEFE, PUTR2, CEVE, RICE, RUID, BERBE, CHVIL4
values_data <- shrub_df$Item.Code
values_valid <- c("CEFE", "CEVE", "PUTR2", "RICE", "RUID", "BERBE", "CHVIL4")
values_check <- values_data %in% values_valid
errors_shrubs <- rbind(errors_shrubs, qaqc(shrub_df, "shrub species", "species", values_check))



```


# Density - Quadrats
1. # transects = 1
2. # quadrats per transect = 4
3. Quad length = 10
4. Quad Width = 10
5. Quad Area = 78.5
6. Transect = 1
7. Quadrat = 1, 2, 3, or 4
8. Status = L
9. Height = 1, 2, 3, 4, 5, 6, or 7

```{r}

# QAQC Checks

# 1. # Transects = 1
values_data <- seedling_headers$Num..Transects
values_valid <- 1
values_check <- values_data == values_valid
errors_seedlings <- qaqc(seedling_headers, "seedling NumTran", "Number of Transects", values_check)

# 2. 4 Quadrats per transect
values_data <- seedling_headers$Num..Quad..Tran.
values_valid <- 4
values_check <- values_data == values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_headers, "seedling QuadPerTran", "Quadrats per transect", values_check))

# 3. Quad length = 10
values_data <- seedling_headers$Quad..Length
values_valid <- 10
values_check <- values_data == values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_headers, "seedling QuadLength", "Quadrat Length", values_check))

# 4. Quad width = 10
values_data <- seedling_headers$Quad..Width
values_valid <- 10
values_check <- values_data == values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_headers, "seedling QuadWidth", "Quadrat Width", values_check))

# 4. Quad Area = 78.5
values_data <- seedling_headers$Quad..Area
values_valid <- 78.5
values_check <- values_data == values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_headers, "seedling QuadArea", "Quadrat Area", values_check))

# 5. Transect = 1
values_data <- seedling_df$Transect
values_valid <- 1
values_check <- values_data == values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_df, "seedling Transect #", "Transect", values_check))

# 6. Quadrat = 1, 2, 3, or 4
values_data <- seedling_df$Quadrat
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_df, "seedling Quadrat #", "Quadrat", values_check))

# 7. Status = L
values_data <- seedling_df$Status
values_valid <- "L"
values_check <- values_data == values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_df, "seedling status", "Status", values_check))

# 8. Height = 1, 2, 3, 4, 5, 6, 7
values_data <- seedling_df$Height
values_valid <- c(1:7)
values_check <- values_data %in% values_valid
errors_seedlings <- rbind(errors_seedlings, qaqc(seedling_df, "seedling height", "Height Class", values_check))




```


## Step 4. Combine error logs and save as .xslx

```{r}
errors_ALL <- rbind(errors_FuelsFine, errors_Fuels1000, errors_FuelsDL, errors_Trees, errors_cover, errors_BS, errors_seedlings)
errors_ALL <- unique(errors_ALL)

write_xlsx(errors_ALL, "O:/resmgmt/Fire Management/Fire Effects/QAQC/Error Logs/RAPASS_2024.xlsx")


```
















