---
title: "6. ROMO RAPASS 2025"
author: "Christina Fossum"
date: "2025-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Date: 09/15/2025
Edited by: Christina Fossum

**test
Summary: QAQC code for 
  # 1. 2024 Fall RX 01yr01 reads (RAP_01, RAP_02, RAP_03, RAP_04, RAP_06, RAP_09, RAP_10) 
  # 2. 2025 Spring 01yr01 reads (RAP_13, RAP_14, RAP_15, RAP_16, PIPS_16, PIPS_08)
  # 3. Pre reads for fall 2025 RX (PIPS_01, PIPS_02, PIPS_05, PIPO_09, PIPO_27, PIPS_14)
  # 4. Thunder MOuntain Plots
  # 5. West Side RAP plots from 2025
  
  ***Note: edit filepath for (2) load data, and fuel constant for fuels queries, and tree species!
  
Contents / Steps:
1. Load packages and functions
2. Load data
3. Surface fuels
  A. Fine
  B. Coarse
  C. Litter/Duff
4. Trees
5. Cover
6. Seedlings
7. Shrubs
8. Combine error logs and save as .xslx
  

##1. Load packages and functions
```{r}

#Load packages
library(tidyverse)
library(dplyr)
library(writexl)
library(knitr)
library(EnvStats)

# Create Blank error log
errors_blank <- data.frame("SavedQuery" = "", "MacroPlot.Name" = "", "Sample.Event.Date" = "", "Error" = "No Error", "Comment" = "")

# QAQC Function
qaqc <- function(data, query, query_message, values_check, values_data = NULL, Comment = "") {
  data_checked <- data %>%
    mutate(SavedQuery = query,
           Error = if (!is.null(values_data)) {
             paste0(query_message, " = ", values_data)
           } else {
             paste0(query_message, " = ", values_check)
           },
           Comment = Comment)

  errors <- data_checked %>%
    filter(!values_check) %>%
    select(any_of(c("SavedQuery", "MacroPlot.Name", "Sample.Event.Date", "Error", "Comment")))

  if (nrow(errors) == 0) {
    errors <- errors_blank %>%
      mutate(SavedQuery = query)
  }

  return(errors)
}


# Header loading function
load_headers <- function(directory_path, pattern) {
  list_files <- list.files(path = directory_path, pattern = pattern, full.names = TRUE)
  data_list <- lapply(list_files, function(file) {
    file_data <- read.csv(file) %>% select(1:10) %>% filter(Visited == "True")
    file_data <- file_data %>% mutate(Comment = "")
  })
  return(do.call(rbind, data_list))
}


# Data Loading Function
load_data <- function(directory_path, pattern) {
  list_files <- list.files(path = directory_path, pattern = pattern, full.names = TRUE)
  data_list <- lapply(list_files, function(file) {
    if (file.info(file)$size == 0) return(NULL)
    file_data <- tryCatch({
      read.csv(file, skip = 3) %>% mutate(Comment = "")
    }, error = function(e) {
      message("Error reading file: ", file, " - skipping this file.")
      return(NULL)
    })
    return(file_data)
  })
  do.call(rbind, data_list[!sapply(data_list, is.null)])
}

```


##2. Load Data
```{r}

#Set Working Directory to QAQC Folder: Session -> Set Working Directory -> Choose Directory

# Set "directory_path" to source file location
directory_path <- "FFI Data - WORKING/2025_Spring2025RX_01yr01/New folder/"
directory_path <- "FFI Data - Working/2025_PICO/"

# load data (Surface Fuels - Fine)
fuelsfine_df <- load_data(directory_path, "_Surface Fuels - Fine.csv$")
fuelsfine_headers <-load_headers(directory_path, "_Surface Fuels - Fine.csv$")

# load data (Surface Fuels - 1000Hr)
fuels1000_df <- load_data(directory_path, "_Surface Fuels - 1000Hr.csv$")
fuels1000_headers <-load_headers(directory_path, "_Surface Fuels - 1000Hr.csv$")

# load data (Surface Fuels - Duff/Litter)
fuelsDL_df <- load_data(directory_path, "_Surface Fuels - Duff_Litter.csv$")
fuelsDL_headers <-load_headers(directory_path, "_Surface Fuels - Duff_Litter.csv$")

# load data (Trees - Individuals)
trees_df <- load_data(directory_path, "_Trees - Individuals \\(metric\\)\\.csv$")

list <- list.files(path = directory_path, pattern = "_Trees - Individuals \\(metric\\)\\.csv$", full.names = TRUE)
trees_headers <- lapply(list, function(file) read.csv(file, nrows = 1))
trees_headers <- do.call(rbind, trees_headers)

# load data (Cover - Points)
cover_df <- load_data(directory_path, "_Cover - Points \\(metric\\)\\.csv$")
cover_headers <-load_headers(directory_path, "_Cover - Points \\(metric\\)\\.csv$")


# load data (Seedlings)
seedling_df <- load_data(directory_path, "_Density - Quadrats \\(metric\\)\\.csv$")
seedling_headers <-load_headers(directory_path, "_Density - Quadrats \\(metric\\)\\.csv$")

# Load data (shrubs)
shrub_df <- load_data(directory_path, "_Density - Belts \\(metric\\)\\.csv$")
shrub_headers <- load_headers(directory_path, "_Density - Belts \\(metric\\)\\.csv$")


```


##3. Surface fuels

#3A. Fine Fuel
Checks:
1. Number of Transects = 2 
2. Transect Length = 65.6 for 1hr, 10hr, 100hr
3. FWD fuel constant = correct
4. Azimuth is 0 for Transect 1, 270 for transect 2
5. 2x transect data per plot
6. Outlier values flagged for 1hr, 10hr, 100hr fuel counts

```{r}
errors_FuelsFine <- tibble()

check_qaqc <- function(data, query, message, condition, values_data = NULL, Comment = "") {
  qaqc(data, query, message, condition, values_data = values_data, Comment = Comment)
}

# 1. Number of Transects = 2
errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    data = fuelsfine_headers,
    query = "Fine Fuels NumTran",
    message = "Number of Transects",
    condition = fuelsfine_headers$Num..Transects == 2
  )
)

# 2. Transect Length = 65.6 for all
values_check <- with(fuelsfine_headers,
  X1.hr.Tran..Len. == 65.6 &
  X10.hr.Tran..Len. == 65.6 &
  X100.hr.Tran..Len. == 65.6
)

values_data <- with(fuelsfine_headers, 
  paste0("X1=", X1.hr.Tran..Len., ", X10=", X10.hr.Tran..Len., ", X100=", X100.hr.Tran..Len.)
)

errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    data = fuelsfine_headers,
    query = "Fine Fuels TranLen",
    message = "Invalid Transect Lengths",
    condition = values_check,
    values_data = values_data
  )
)



# 3. Fuel Constant = "Default"
errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    fuelsfine_df,
    "Fuel Fine FuelConst",
    "Fuel Constant",
    fuelsfine_df$FWDFuConSt == "Default"
  )
)

# 4. Azimuth Check
azimuth_check <- with(fuelsfine_df,
  (Transect == 1 & Azimuth == 0) |
  (Transect == 2 & Azimuth == 270)
)
errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    fuelsfine_df,
    "Fine Fuel Azimuth",
    "Correct Azimuth",
    azimuth_check
  )
)

# 5. Hits Check: Ensure 2 hits per plot
fuelsfine_df <- fuelsfine_df %>%
  mutate(hits = 1) %>%
  group_by(MacroPlot.Name, Sample.Event.Date) %>%
  mutate(sumhits = sum(hits)) %>%
  ungroup()

errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    fuelsfine_df,
    "Fuel Fine Hits",
    "Number of Hits",
    fuelsfine_df$sumhits == 2
  )
)

# 6. Z-score Outlier Checks (for 1hr, 10hr, 100hr)
detect_outliers_zscore <- function(x) {
  z <- (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
  !is.na(z) & abs(z) <= 3  # TRUE means "not an outlier"
}

# 1hr
outlier_flag <- !detect_outliers_zscore(fuelsfine_df$X1.Hr.Count)  # TRUE = outlier
values_data <- ifelse(
  outlier_flag,
  paste0("Outlier Value: ", fuelsfine_df$X1.Hr.Count),
  NA_character_
)

errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    data = fuelsfine_df,
    query = "Fine Fuels x1hr value",
    message = "One Hour Count Outlier (Z-score)",
    condition = !outlier_flag,
    values_data = values_data
  )
)

# 10hr
outlier_flag_10 <- !detect_outliers_zscore(fuelsfine_df$X10.Hr.Count)
values_data_10 <- ifelse(
  outlier_flag_10,
  paste0("Outlier Value: ", fuelsfine_df$X10.Hr.Count),
  NA_character_
)
errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    fuelsfine_df,
    "Fine Fuels x10hr value",
    "10 Hour Count Outlier (Z-score)",
    condition = !outlier_flag_10,
    values_data = values_data_10
  )
)

# 100hr
outlier_flag_100 <- !detect_outliers_zscore(fuelsfine_df$X100.Hr.Count)
values_data_100 <- ifelse(
  outlier_flag_100,
  paste0("Outlier Value: ", fuelsfine_df$X100.Hr.Count),
  NA_character_
)
errors_FuelsFine <- bind_rows(errors_FuelsFine,
  check_qaqc(
    fuelsfine_df,
    "Fine Fuels x100hr value",
    "100 Hour Count Outlier (Z-score)",
    condition = !outlier_flag_100,
    values_data = values_data_100
  )
)



```


#3B. Coarse Fuel
Checks: 
1. Number of transects = 2
2. Transect length = 65.6
3. Correct CWD Fuel Constant
4. Decay Class is either 3 or 4
5. Outliers flagged for diameter

```{r}

# 1. Number of Transects = 2
values_data <- fuels1000_headers$Num..Transects
values_valid <- 2
values_check <- values_data == values_valid
errors_Fuels1000 <- qaqc(fuels1000_headers, 
                         "1000hr Fuels NumTran", 
                         "Number of Transects", 
                         values_check,
                         values_data = values_data)

# 2. Transect Length = 65.6
values_data <- fuels1000_headers$Tran..Length
values_valid <- 65.6
values_check <- values_data == values_valid
errors_Fuels1000 <- rbind(errors_Fuels1000, 
                         qaqc(fuels1000_headers, 
                              "x1000 Fuels TranLen", 
                              "Transect Length", 
                              values_check,
                              values_data = values_data))

# 3. Check Fuel Constant
values_data <- fuels1000_df$CWDFuConSt
values_valid <- "Default"
values_check <- values_data == values_valid
errors_Fuels1000 <- rbind(errors_Fuels1000, 
                         qaqc(fuels1000_df, 
                              "x1000 CWD FuelConst", 
                              "Fuel Constant", 
                              values_check,
                              values_data = values_data))

# 4. Decay class 3 or 4
valid_decay_classes <- c(3, 4)
values_data <- fuels1000_df$Decay.Class
values_check <- values_data %in% valid_decay_classes
errors_Fuels1000 <- rbind(errors_Fuels1000, 
                         qaqc(fuels1000_df, 
                              "x1000 decay class", 
                              "Decay Class (3 or 4)", 
                              values_check,
                              values_data = values_data))

# 5. Outliers flagged for Diameter

# Re-using your detect_outliers_zscore function
outliers_z <- detect_outliers_zscore(fuels1000_df$Diameter)
values_check_z <- !outliers_z

# For error messages, show the actual diameter values where outliers flagged
values_data_z <- ifelse(outliers_z, 
                        paste0("Outlier Diameter: ", fuels1000_df$Diameter), 
                        NA_character_)

errors_Fuels1000 <- rbind(errors_Fuels1000, 
                         qaqc(fuels1000_df, 
                              "x1000 Dia value", 
                              "Diameter Outlier (Z-score)", 
                              values_check_z,
                              values_data = values_data_z))
```

#3C. Litter/Duff
Checks:
1. # of transects = 2
2. Correct Sample Locations 
3. # of samples per transect = 10
4. Correct DL Fuel Constant
5. Outlier values flagged for duff, litter

```{r}

# 1. Number of Transects = 2
values_data <- fuelsDL_headers$Num..Transects
values_valid <- 2
values_check <- values_data == values_valid
errors_FuelsDL <- qaqc(fuelsDL_headers, 
                      "DL NumTran", 
                      "Number of Transects", 
                      values_check, 
                      values_data = values_data)

# 2. Correct Sample Locations
values_data <- fuelsDL_df$Samp..Loc.
values_valid <- c(1, 3, 5, 7, 9, 11, 13, 15, 17, 19)
values_check <- values_data %in% values_valid
errors_FuelsDL <- rbind(errors_FuelsDL,
                       qaqc(fuelsDL_df, 
                            "DL Sample Location", 
                            "Invalid Sample Location", 
                            values_check, 
                            values_data = values_data))

# 3. # samples per transect = 10
fuelsDL_df <- fuelsDL_df %>%
  mutate(hits = 1) %>%
  group_by(MacroPlot.Name, Sample.Event.Date, Transect) %>%
  mutate(sumhits = sum(hits)) %>%
  ungroup()

values_data <- fuelsDL_df$sumhits
values_valid <- 10
values_check <- values_data == values_valid
errors_FuelsDL <- rbind(errors_FuelsDL,
                       qaqc(fuelsDL_df, 
                            "LitterDuff Hits Per Transect", 
                            "Hits Per Transect", 
                            values_check, 
                            values_data = values_data))

# 4. Correct DL Fuel Constant
values_data <- fuelsDL_df$DLFuConSt
values_valid <- "Default"
values_check <- values_data == values_valid
errors_FuelsDL <- rbind(errors_FuelsDL,
                       qaqc(fuelsDL_df, 
                            "DL FuelConst", 
                            "Fuel Constant", 
                            values_check, 
                            values_data = values_data))

# 5. Outliers flagged (Litter.Depth, Duff.Depth, Fuelbed.Depth)
detect_outliers_zscore <- function(data) {
  z_scores <- (data - mean(data, na.rm = TRUE)) / sd(data, na.rm = TRUE)
  abs(z_scores) > 3
}

# Litter.Depth outliers
outliers_litt_z <- detect_outliers_zscore(fuelsDL_df$Litter.Depth)
values_check_litt_z <- !outliers_litt_z
values_data_litt <- ifelse(outliers_litt_z, 
                          paste0("Outlier Litter Depth: ", fuelsDL_df$Litter.Depth),
                          NA_character_)
errors_FuelsDL <- rbind(errors_FuelsDL,
                       qaqc(fuelsDL_df,
                            "Litter Outliers",
                            "Litter Depth (z-score)",
                            values_check_litt_z,
                            values_data = values_data_litt))

# Duff.Depth outliers
outliers_duff_z <- detect_outliers_zscore(fuelsDL_df$Duff.Depth)
values_check_duff_z <- !outliers_duff_z
values_data_duff <- ifelse(outliers_duff_z,
                          paste0("Outlier Duff Depth: ", fuelsDL_df$Duff.Depth),
                          NA_character_)
errors_FuelsDL <- rbind(errors_FuelsDL,
                       qaqc(fuelsDL_df,
                            "Duff Outliers",
                            "Duff Depth (z-score)",
                            values_check_duff_z,
                            values_data = values_data_duff))



```


##4. Trees

Checks: 
1. Subplot = 1 
2. Quarter = 1, 2, 3, or 4
3. Correct Plot Area
4. Correct Snag Plot Area
5. Correct Break Pt. Diameter
6. Species is appropriate (flags for outliers)
7. Correct Crown Class 
8. Outlier values flagged for DBH, height, CBH
9. Reasonable damage code




```{r}

# QAQC Checks

# 1. Subplot = 1 
values_data <- trees_df$Subplot.Frac.
valid_values <- 1
values_check <- values_data %in% valid_values
errors_Trees <- qaqc(
  trees_df,
  "Subplot Fraction",
  "Subplot must be 1",
  values_check,
  values_data = values_data
)

# 2. Quarter = 1, 2, 3, 4
values_data <- trees_df$Quarter
valid_values <- c(1, 2, 3, 4)
values_check <- values_data %in% valid_values
errors_Trees <- rbind(errors_Trees,
  qaqc(
    trees_df,
    "Tree Quarter",
    "Quarter must be 1-4",
    values_check,
    values_data = values_data
  )
)

# 3. Plot Area
values_data <- trees_headers$Plot.Area
values_check <- values_data == 0.032
errors_Trees <- rbind(errors_Trees,
  qaqc(
    trees_headers,
    "Tree Plot Area",
    "Plot Area must be 0.032",
    values_check,
    values_data = values_data
  )
)

# 4. Snag Plot Area
values_data <- trees_headers$Snag.Plot.Area
values_check <- values_data == 0.032
errors_Trees <- rbind(errors_Trees,
  qaqc(
    trees_headers,
    "Snag Plot Area",
    "Snag Plot Area must be 0.032",
    values_check,
    values_data = values_data
  )
)

# 5. Break Point Diameter
values_data <- trees_headers$Break.Pnt..Dia.
values_check <- values_data == 15
errors_Trees <- rbind(errors_Trees,
  qaqc(
    trees_headers,
    "Tree Break Point Diameter",
    "Break Point Diameter must be 15",
    values_check,
    values_data = values_data
  )
)


# 6. Reasonable tree species

species_freq <- table(trees_df$Species)
rare_species <- names(species_freq[species_freq < 5])

outliers_species <- trees_df$Species %in% rare_species
values_check <- !outliers_species
values_data <- ifelse(outliers_species, paste0("Outlier Species: ", trees_df$Species), NA)

errors_Trees <- rbind(
  errors_Trees,
  qaqc(trees_df, "Species Outliers", "Rare species (frequency < 5)", values_check, values_data)
)



# 7. Crown Class logic 
values_check <- case_when(
  trees_df$Status == "D" & trees_df$Crown.Class %in% c("BAD", "BBD", "CS", "LBS", "RS", "DD") ~ TRUE,
  trees_df$Status == "L" & trees_df$Crown.Class %in% c("I", "O", "SC", "D", "C") ~ TRUE,
  TRUE ~ FALSE
)

values_data <- paste0("Status: ", trees_df$Status, ", Crown.Class: ", trees_df$Crown.Class)

errors_Trees <- rbind(
  errors_Trees,
  qaqc(
    trees_df,
    query = "Tree CPC",
    query_message = "Invalid Crown Class for Status",
    values_check = values_check,
    values_data = values_data
  )
)

# 8. Outlier detection for DBH, CBH, Height

detect_outliers_zscore <- function(x) {
  z <- (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)
  abs(z) > 3
}

# DBH
outliers_dbh <- detect_outliers_zscore(trees_df$DBH)
values_check <- !outliers_dbh
values_data <- ifelse(outliers_dbh, paste0("Outlier DBH: ", trees_df$DBH), NA)
errors_Trees <- rbind(errors_Trees,
  qaqc(trees_df, "DBH Outliers", "DBH outlier (z-score)", values_check, values_data)
)


# Height
outliers_height <- detect_outliers_zscore(trees_df$Height)
values_check <- !outliers_height
values_data <- ifelse(outliers_height, paste0("Outlier Height: ", trees_df$Height), NA)
errors_Trees <- rbind(errors_Trees,
  qaqc(trees_df, "Height Outliers", "Height outlier (z-score)", values_check, values_data)
)

#CBH
outliers_cbh <- detect_outliers_zscore(trees_df$Live.Crown.Base.Ht.)
values_check <- !outliers_cbh
values_data <- ifelse(outliers_cbh, paste0("Outlier CBH: ", trees_df$Live.Crown.Base.Ht.), NA)
errors_Trees <- rbind(errors_Trees,
  qaqc(trees_df, "CBH Outliers", "CBH outlier (z-score)", values_check, values_data)
)


# 9. Reasonable Damage code (Check for outliers)

all_damages <- unlist(trees_df[, c("Damage.1", "Damage.2", "Damage.3", "Damage.4", "Damage.5")])
all_damages <- all_damages[!is.na(all_damages)]
damage_freq <- table(all_damages)
rare_damages <- names(damage_freq[damage_freq < 2])

damage_cols <- c("Damage.1", "Damage.2", "Damage.3", "Damage.4", "Damage.5")
outliers_damage <- apply(trees_df[, damage_cols], 1, function(row) {
  any(row %in% rare_damages, na.rm = TRUE)
})

values_data <- apply(trees_df[, damage_cols], 1, function(row) {
  rare_found <- row[row %in% rare_damages]
  if (length(rare_found) > 0) {
    paste("Outlier Damage Code(s):", paste(unique(rare_found), collapse = ", "))
  } else {
    NA
  }
})

values_check <- !outliers_damage

errors_Trees <- rbind(
  errors_Trees,
  qaqc(trees_df, "Damage Code Outliers", "Rare damage codes (frequency < 2)", values_check, values_data)
)




```


##5. Cover

Checks: 
1. # of transects = 2
2. Transect Length = 20m
3. Points per transect = 20
4. 'Point' = 'Tape' 
5. NumPts = 20
6. 1st hit has height value only if a plant (no other height values)
7. Outlier values flagged for plant height
8. Outlier values flagged for cover species

```{r}

# QAQC Checks 

# 1. Number of Transects = 2
values_data <- cover_headers$Num..Transects
values_check <- values_data == 2
errors_cover <- qaqc(
  data = cover_headers,
  query = "Cover NumTran",
  query_message = "Number of Transects",
  values_check = values_check,
  values_data = values_data
)

# 2. Transect Length = 20m
values_data <- cover_headers$Tran..Length
values_check <- values_data == 20
errors_cover <- bind_rows(errors_cover, qaqc(
  data = cover_headers,
  query = "Cover Tran Length",
  query_message = "Transect Length",
  values_check = values_check,
  values_data = values_data
))

# 3. Points per Transect = 20
values_data <- cover_headers$Num..Pts..Tran.
values_check <- values_data == 20
errors_cover <- bind_rows(errors_cover, qaqc(
  data = cover_headers,
  query = "Cover Points Per Transect",
  query_message = "Points per Transect",
  values_check = values_check,
  values_data = values_data
))

# 4. Point == Tape
values_data <- paste0("Point: ", cover_df$Point, " | Tape: ", cover_df$Tape)
values_check <- cover_df$Point == cover_df$Tape
errors_cover <- bind_rows(errors_cover, qaqc(
  data = cover_df,
  query = "Cover Point/Tape Match",
  query_message = "Point matches Tape",
  values_check = values_check,
  values_data = values_data
))

# 5. Number of unique Points per Transect = 20
cover_df <- cover_df %>%
  group_by(MacroPlot.Name, Sample.Event.Date, Transect) %>%
  mutate(NumPts = n_distinct(Point)) %>%
  ungroup()

values_data <- cover_df$NumPts
values_check <- values_data == 20
errors_cover <- bind_rows(errors_cover, qaqc(
  data = cover_df,
  query = "Cover Obs Count",
  query_message = "Points per Transect = 20",
  values_check = values_check,
  values_data = values_data
))

# 6. Height logic: only plant species (on first hit) have height values
values_check <- case_when(
  cover_df$Order == 1 & !(cover_df$Species %in% c("LITT1", "WOOD1", "BARE1", "ROCK1")) & !is.na(cover_df$Height) ~ TRUE,
  cover_df$Order == 1 & (cover_df$Species %in% c("LITT1", "WOOD1", "BARE1", "ROCK1")) & is.na(cover_df$Height) ~ TRUE,
  cover_df$Order != 1 & is.na(cover_df$Height) ~ TRUE,
  TRUE ~ FALSE
)

values_data <- paste0("Order: ", cover_df$Order,
                      ", Species: ", cover_df$Species,
                      ", Height: ", cover_df$Height)

errors_cover <- bind_rows(errors_cover, qaqc(
  data = cover_df,
  query = "Cover Height Logic",
  query_message = "Height present only for valid species (order 1)",
  values_check = values_check,
  values_data = values_data
))

# 7. Outliers - Height
outliers_z <- detect_outliers_zscore(cover_df$Height)
values_check <- !outliers_z
values_data <- cover_df$Height

errors_cover <- bind_rows(errors_cover, qaqc(
  data = cover_df,
  query = "Height Outliers",
  query_message = "Height outlier (Z-score)",
  values_check = values_check,
  values_data = values_data
))

# 8. Species with <2 records flagged
cover_df <- cover_df %>%
  group_by(Species) %>%
  mutate(SpeciesFreq = n()) %>%
  ungroup()

values_check <- cover_df$SpeciesFreq >= 2
values_data <- paste0("Species: ", cover_df$Species, ", Freq: ", cover_df$SpeciesFreq)

errors_cover <- bind_rows(errors_cover, qaqc(
  data = cover_df,
  query = "Rare Species",
  query_message = "Species occurs < 2 times",
  values_check = values_check,
  values_data = values_data
))
```


##6. Seedlings
1. # transects = 1
2. # quadrats per transect = 4
3. Quad length = 10
4. Quad Width = 10
5. Quad Area = 78.5
6. Transect = 1
7. Quadrat = 1, 2, 3, or 4
8. Status = L or D
9. Height = 0.15, 0.3, 0.6, 1, 2, 3, 4

```{r}

# 1. Number of Transects = 1
values_data <- seedling_headers$Num..Transects
values_check <- values_data == 1
errors_seedlings <- qaqc(
  data = seedling_headers,
  query = "Seedling NumTran",
  query_message = "Number of Transects",
  values_check = values_check,
  values_data = values_data
)

# 2. Quadrats per Transect = 4
values_data <- seedling_headers$Num..Quad..Tran.
values_check <- values_data == 4
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_headers,
  query = "Seedling QuadPerTran",
  query_message = "Quadrats per Transect",
  values_check = values_check,
  values_data = values_data
))

# 3. Quadrat Length = 10
values_data <- seedling_headers$Quad..Length
values_check <- values_data == 10
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_headers,
  query = "Seedling QuadLength",
  query_message = "Quadrat Length",
  values_check = values_check,
  values_data = values_data
))

# 4. Quadrat Width = 10
values_data <- seedling_headers$Quad..Width
values_check <- values_data == 10
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_headers,
  query = "Seedling QuadWidth",
  query_message = "Quadrat Width",
  values_check = values_check,
  values_data = values_data
))

# 5. Quadrat Area = 78.5
values_data <- seedling_headers$Quad..Area
values_check <- values_data == 78.5
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_headers,
  query = "Seedling QuadArea",
  query_message = "Quadrat Area",
  values_check = values_check,
  values_data = values_data
))

# 6. Transect = 1
values_data <- seedling_df$Transect
values_check <- values_data == 1
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_df,
  query = "Seedling Transect",
  query_message = "Transect number must be 1",
  values_check = values_check,
  values_data = values_data
))

# 7. Quadrat = 1, 2, 3, 4
values_data <- seedling_df$Quadrat
values_check <- values_data %in% c(1, 2, 3, 4)
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_df,
  query = "Seedling Quadrat #",
  query_message = "Quadrat must be 1–4",
  values_check = values_check,
  values_data = values_data
))

# 8. Status = L or D
values_data <- seedling_df$Status
valid_values <- c("L", "D")
values_check <- values_data %in% valid_values
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_df,
  query = "Seedling Status",
  query_message = "Status must be L or D",
  values_check = values_check,
  values_data = values_data
))

# 9. Height in valid classes
values_data <- seedling_df$Height
valid_values <- c(0.15, 0.3, 0.6, 1, 2, 3, 4)
values_check <- values_data %in% valid_values
errors_seedlings <- bind_rows(errors_seedlings, qaqc(
  data = seedling_df,
  query = "Seedling Height",
  query_message = "Height class invalid",
  values_check = values_check,
  values_data = values_data
))


```


##7. Shrubs 
Checks:
1. # of transects = 2
2. # of sub-belts = 4
3. Transect length = 20m
4. Transect Width = 2m
5. Transect Area = 40m
6. Sub-belt = 1, 2, 3, 4
7. Transect = 1 or 2
8. Age Class = I, M, or R
9. Status = L or D
10. Species = CEFE, PUTR2, CEVE, RICE, RUID, BERBE, CHVIL4

```{r}

# QAQC Checks

# 1. Number of Transects = 2
values_data <- shrub_headers$Num..Transects
values_check <- values_data == 2
errors_shrubs <- qaqc(
  data = shrub_headers,
  query = "Shrub NumTran",
  query_message = "Number of Transects",
  values_check = values_check,
  values_data = values_data
)

# 2. Subbelts per Transect = 4
values_data <- shrub_headers$Num..Subbelts
values_check <- values_data == 4
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_headers, "Shrub NumSubbelts", "Subbelts per transect", values_check, values_data
))

# 3. Transect Length = 20m
values_data <- shrub_headers$Tran..Length
values_check <- values_data == 20
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_headers, "Shrub TranLength", "Transect Length", values_check, values_data
))

# 4. Transect Width = 2m
values_data <- shrub_headers$Tran..Width
values_check <- values_data == 2
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_headers, "Shrub TranWidth", "Transect Width", values_check, values_data
))

# 5. Transect Area = 40m²
values_data <- shrub_headers$Tran..Area
values_check <- values_data == 40
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_headers, "Shrub TranArea", "Transect Area", values_check, values_data
))

# 6. Subbelt = 1 to 4
values_data <- shrub_df$Subbelt
values_check <- values_data %in% 1:4
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_df, "Shrub Subbelt", "Subbelt must be 1–4", values_check, values_data
))

# 7. Transect = 1 or 2
values_data <- shrub_df$Transect
values_check <- values_data %in% 1:2
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_df, "Shrub Transect", "Transect must be 1 or 2", values_check, values_data
))

# 8. Age Class = I, M, or R
values_data <- shrub_df$Age.Class
values_check <- values_data %in% c("I", "M", "R")
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_df, "Shrub AgeClass", "Age Class must be I, M, or R", values_check, values_data
))

# 9. Status = L or D
values_data <- shrub_df$Status
values_check <- values_data %in% c("L", "D")
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_df, "Shrub Status", "Status must be L or D", values_check, values_data
))


# 10. Species = Approved codes
values_data <- shrub_df$Item.Code
valid_species <- c("CEFE", "CEVE", "PUTR2", "RICE", "RUID", "BERBE", "CHVIL4", "ROWO")
values_check <- values_data %in% valid_species
errors_shrubs <- bind_rows(errors_shrubs, qaqc(
  shrub_df, "Shrub Species", "Unrecognized Species Code", values_check, values_data
))


```


## Step 8. Combine error logs and save as .xslx

```{r}
errors_ALL <- rbind(errors_FuelsFine, errors_Fuels1000, errors_FuelsDL, errors_Trees, errors_cover, errors_seedlings, errors_shrubs)
errors_ALL <- unique(errors_ALL)

write_xlsx(errors_ALL,"Error Logs/2025_2025SpringRX.xlsx")


```












