---
title: "DINO_ErrorCheck_2024"
author: "Christina Fossum"
date: "2025-01-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Overview
Created by Christina Fossum, 12/31/2024

This R-Markdown file contains code to perform QA/QC checks on the DINO Pearl Park plots. Each protocol is Analyzed individually.

Protocols:
1. Surface Fuels - Fine 
2. Surface Fuels - 1000Hr 
3. Surface Fuels - Duff_Litter 
4. Trees - Individuals (metric)
5. Cover - Points (metric)
6. Density - Belts (metric)
7. Density - Quadrats (metric)

* Note: This Document performs QAQC on the FMH plot data collected in 2024. You may adapt this document to perform QAQC on whichever data/plots but make sure to Save As a new document before making changes. 

All Raw FFI export files are located in:
"C:/Users/cfossum/OneDrive - DOI/ROMO Fire Effects 2024/Data Management/2024 Data/2024 FFI Exports - RAW/FMH FFI export 2024/"

The objective of this document is to produce an error log to easily identify and then correct data in FFI. 


# Step 0 - Install / Load packages
```{r}
# 1. Install packages (you must do this first before you load them. Once you have installed, you can just load in the future)
#packages <- c("tidyverse", "dplyr", "writexl", "knitr", "EnvStats")
#install.packages(packages)

# 1B. gov. computers might block you installing packages. If this happens, run the following line of code and then re-start R, then try step 1 again
#system2("setx", args = c("R_LIBCURL_SSL_REVOKE_BEST_EFFORT", "TRUE"))

# 3. Load packages (This is the only step you need to run once you have installed)
library(tidyverse)
library(dplyr)
library(writexl)
library(knitr)
library(EnvStats)

```

# Create QAQC Error Log
```{r}

# Create Blank error log
errors_blank <- data.frame("SavedQuery" = "", "MacroPlot.Name" = "",
                           "Sample.Event.Date" = "","Error" = "No Error", 
                           "Comment" = "")

# 3. Create QAQC Function
qaqc <- function(data, query, query_message, values_check, Comment) {
  # Identify errors in the data based on the provided check condition
  errors <- data %>%   
    # Create new columns: "SavedQuery" with the query parameter and "Error"
    # which concatenates the query_message and values_data
    mutate("SavedQuery" = query, 
           "Error" = paste(query_message, "=", values_data)) %>%
    # Filter the data to find rows where values_check condition is TRUE
    # use replace_na to treat NA values as TRUE (which means they pass the check)
    filter(!(values_check) %>% replace_na(TRUE)) %>%   
    # Select specific columns to include the errors data frame
    select("SavedQuery", "MacroPlot.Name", "Sample.Event.Date", "Error", "Comment")
  # If no errors are identified, add "No Errors" to existing data frame 
  if (nrow(errors) == 0) {   
    errors <- errors_blank %>%     
      mutate(SavedQuery = query) 
  } else {   
    # if there are errors, retain the errors data frame as is
    errors <- errors
  }
}

```


#################################################################################################################################

## Cover

# Load data & Create Error Log
1. Number of transects = 2
2. Transect length = 20
3. Points per transect = 20
4. Point = Tape
5. Each transect should have at least 1 observation for each point
6. Status should be X for bare, litter, rock, wood; otherwise should be L or D
8. Outliers flagged for species 

```{r}
directory_path <- "C:/Users/cfossum/OneDrive - DOI/ROMO Fire Effects 2024/Data Management/2024 Data/2024 FFI Exports - RAW/DINO PearlPark export ALLYEARS/"

list <- list.files(path = directory_path, pattern = "_Cover - Points \\(metric\\)\\.csv$", full.names = TRUE)
num <- 1:length(list)
xfunction <- function(x){
  first_file <- list[x]
  file <- read.csv(first_file, skip = 3)
  file2 <- read.csv(first_file)
  file2 <- file2 %>% select(1:10) %>% filter(Visited == "True")
  file3 <- left_join(file, file2)
}
list <- lapply(num, xfunction)
cover_df <- do.call(rbind, list)

#########################################################################################

## 1. Number of Transects = 1
data <- cover_df
query <- "cover NumTran"
query_message <- "number of transects"
values_data <- cover_df$Num..Transects
values_valid <- 1
values_check <- values_valid == values_data
#add to error log
errors_cover <- qaqc(data, query, query_message, values_check)

# 2. Transect length = 100
data <- cover_df
query <- "cover TranLength"
query_message <- "transect length"
values_data <- cover_df$Tran..Length
values_valid <- 100
values_check <- values_valid == values_data
#add to error log
errors_cover <- rbind(errors_cover, qaqc(data, query, query_message, values_check))

# 3. Points per transect = 200
data <- cover_df
query <- "cover TranPts"
query_message <- "points per transect"
values_data <- cover_df$Num..Pts..Tran.
values_valid <- 200
values_check <- values_valid == values_data
#add to error log
errors_cover <- rbind(errors_cover, qaqc(data, query, query_message, values_check))

errors_cover <- unique(errors_cover)

# 4. Each transect should have at least 1 observation at each point
cover_df <- cover_df %>% group_by(MacroPlot.Name, Sample.Event.Date, Transect) %>% mutate(NumPts = n_distinct(Point)) %>% mutate(NumTapePts = n_distinct(Tape))%>% ungroup()

data <- cover_df
query <- "cover observation number"
query_name <- "correct # of observations"
values_data <- cover_df$NumPts
values_valid <- 100
values_check <- values_data == values_valid
#add to error log
errors_cover <- rbind(errors_cover, qaqc(data, query, query_message, values_check))

# 5. Each transect should have at least 1 observation at each tape point

data <- cover_df
query <- "Cover Tape"
query_name <- "Cover tape location"
values_data <- cover_df$NumTapePts
values_valid <- 100
values_check <- values_data == values_valid
#add to error log
errors_cover <- rbind(errors_cover, qaqc(data, query, query_message, values_check))

errors_cover <- unique(errors_cover)

# 6. Status should be X for only litt, wood, bare, rock

data <- cover_df
query <- "cover status"
query_message <- "status"
values_data <- cover_df$Status
values_valid <- case_when(cover_df$Species %in% c("LITT", "WOOD", "BARE", "ROCK") & cover_df$Status == "X" ~TRUE, !(cover_df$Species %in% c("LITT", "WOOD", "BARE", "ROCK")) & cover_df$Status %in% c("L", "D") ~ TRUE, TRUE ~ FALSE )
values_check <- values_valid
#add to error log
errors_cover <- rbind(errors_cover, qaqc(data, query, query_message, values_check))

errors_cover <- unique(errors_cover)


# 8. Check Species
cover_df <- cover_df %>% group_by(Species) %>% mutate(SpeciesFreq = n()) %>% ungroup()
query <- "Species reasonable"
query_message <- "species"
values_data <- cover_df$Species
valid_values <- case_when(cover_df$SpeciesFreq >= 2 ~ TRUE, TRUE ~ FALSE)
values_check <- valid_values
#Add to error log
errors_cover <- rbind(errors_cover, qaqc(data, query, query_message, values_check))




write_xlsx(errors_cover, "Error Logs/DINO_errors.xlsx")




```




